Black swans in ecological time series\\Black swans in population dynamics\\Heavy-tailed ecological time series\\Black swans in ecological dynamics

### 2014-10-06 figs.tex
Substantial additions to text, supp text, and figures

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.42\textwidth]{nu-coefs-2.pdf}
\caption{
Estimates of how heavy-tailed population dynamics are for \nuCoefPopN\ populations of birds, mammals, insects, and fishes.
Small values of $\nu$ ($\lesssim 10$) suggest heavy-tailed dynamics; larger values of $\nu$ suggest approximately normal-tailed dynamics.
Vertical points and line segments represent posterior medians and 50\% / 90\% credible intervals for individual populations.
Inset plots show probability that $\nu < 10$ (probability of heavy tails) for populations arranged by taxonomic order.
Taxonomic orders are sorted by decreasing mean p($\nu < 10$).
Taxonomic orders with three or fewer populations in panel a are omitted for space.
Red and orange colours highlight populations with a high probability of heavy-tailed dynamics.
}
\label{fig:nu-coefs}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{correlates.pdf}
\caption{
Potential correlates of heavy-tailed behaviour (indicated by low $\nu$ values). Shown are (a--c) parameters from the Gompertz heavy-tailed population model, (b) number of time steps, (c) body length, and (d) lifespan.
For the Gompertz parameters, $\sigma$ refers to the scale parameter of the Student-t process-noise distribution, $\lambda$ refers to the population growth rate at an abundance of one, $b$ refers to the density dependence parameter ($1$ is maximally density independent, $0$ is maximally density dependent, and $<0$ is inversely density dependent).
For parameter values, points and lines represent medians and 50\% credible intervals of the posterior.
Circles representing four populations are filled with white and represent a shark, crustaceans, and one gastropod.
}
\label{fig:correlates}
\end{center}
\end{figure}

>>>>

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.44\textwidth]{nu-coefs-2.pdf}
\caption{
Estimates of how heavy-tailed population dynamics are for \nuCoefPopN\ populations of birds, mammals, insects, and fishes.
Small values of $\nu$ ($\lesssim 10$) suggest heavy-tailed dynamics; larger values of $\nu$ suggest approximately normal-tailed dynamics.
Vertical points and line segments represent posterior medians and 50\% / 90\% credible intervals for individual populations.
Inset plots show probability that $\nu < 10$ (probability of heavy tails) for populations arranged by taxonomic order.
Taxonomic orders are sorted by decreasing mean p($\nu < 10$).
Taxonomic orders with three or fewer populations in panel a are omitted for space.
Red and orange colours highlight populations with a high probability of heavy-tailed dynamics.
}
\label{fig:nu-coefs}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=0.9\textwidth]{correlates.pdf}
\includegraphics[width=0.9\textwidth]{correlates-p10.pdf}
\caption{
Potential correlates of heavy-tailed behaviour (indicated by a high probability that $\nu < 10$).
Shown are (a--c) parameters from the Gompertz heavy-tailed population model, (b) number of time steps, (c) body length, and (d) lifespan.
For the Gompertz parameters, $\sigma$ refers to the scale parameter of the Student-t process-noise distribution, $\lambda$ refers to the population growth rate at an abundance of one, $b$ refers to the density dependence parameter ($1$ is maximally density independent, $0$ is maximally density dependent, and $<0$ is inversely density dependent).
%For parameter values, points and lines represent medians and 50\% credible intervals of the posterior.
Circles representing a limited number of sharks, crustaceans, and gastropods are filled in white.
Median and 90\% credible interval posterior predictions of a beta regression multilevel model are shown in panels a and d where there was a high probability the slope coefficient was different from zero (Supporting Material).
}
\label{fig:correlates}
\end{center}
\end{figure}

### 2014-10-06 supp-figs.tex
Substantial additions to text, supp text, and figures

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{log10-assumed.pdf} \caption{Populations in which the abundance was specified as or assumed to be recorded as log10 values. The numbers at the front of each panel label are the GPDD IDs.}
    \label{fig:log10-assumed}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{priors-gomp-base.pdf} \caption{Probability
  density of the Bayesian priors for the base Gompertz model. From left to right: (1) per capita growth
  rate at $\ln$(abundance) = $0$: $\lambda \sim \mathrm{Normal}(0, 10^2)$; (2) scale
  parameter of t-distribution process noise: $\sigma \sim \mathrm{Half\mhyphen
    Cauchy} (2.5)$; (3) t-distribution degrees of freedom parameter: $\nu \sim
  \mathrm{Exponential}(0.01)$. (4) AR1 correlation coefficient of residuals: $\phi \sim \mathrm{Truncated \mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. Not shown is $b$, the density dependence
  parameter: $b \sim \mathrm{Uniform}(-1, 2)$.}
  \label{fig:priors}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{admb-coefs}
\includegraphics[width=0.8\textwidth]{length-vs-prob-tails}
\caption{Potential correlates of heavy-tailed dynamics.
  (Top panels) Dots, thick lines, and thin lines represent fixed-effect estimates and their 50\% and 95\% confidence intervals for the covariates listed on the left: $\log \sigma$, $b$, $\lambda$, $\log$(lifespan), $\log$(data set length). All covariates were standardized by subtracting their mean and dividing by their standard deviation.
  The beta distribution model (left panel) uses p$(\nu < 10)$ as the response (probability of heavy tails).
  The binomial distribution model (right panel) uses p$(\nu < 10) > 0.5$ vs.\ p$(\nu < 10) \le 0.5$ as the response (heavy vs.\ not heavy tails).
  The models were fitted as mixed effects models with beta or binomial distributions, logit links, and multilevel intercepts for taxonomic class and order nested within class using the \texttt{glmmADMB} \texttt{R} package \citep{fournier2012,glmmadmb}.
  (Bottom panels) Model predictions for the strongest covariate --- time series length --- with all other predictors set to zero (their mean value when standardized). Dots represent observations for individual populations. Lines and shaded regions represent estimates and 50\% and 95\% confidence intervals.
}
    \label{fig:correlate-coefs}
\end{center}
\end{figure}

@@

>>>>

%\begin{figure}[htbp]
%\begin{center}
%\includegraphics[width=\textwidth]{log10-assumed.pdf} \caption{Populations in which the abundance was specified as or assumed to be recorded as log10 values. The numbers at the front of each panel label are the GPDD IDs.}
    %\label{fig:log10-assumed}
%\end{center}
%\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{priors-gomp-base.pdf} \caption{Probability
  density of the Bayesian priors for the Gompertz models. From left to right: (1) per capita growth
  rate at $\ln$(abundance) = $0$: $\lambda \sim \mathrm{Normal}(0, 10^2)$; (2) scale
  parameter of t-distribution process noise: $\sigma \sim \mathrm{Half\mhyphen
    Cauchy} (0, 2.5)$; (3) t-distribution degrees of freedom parameter: $\nu \sim
  \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2)$; (4) AR1 correlation coefficient of residuals: $\phi \sim \mathrm{Truncated \mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. Not shown is $b$, the density dependence parameter: $b \sim \mathrm{Uniform}(-1, 2)$.}
  \label{fig:priors}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=0.7\textwidth]{admb-coefs}
%\includegraphics[width=0.8\textwidth]{length-vs-prob-tails}
\includegraphics[width=0.5\textwidth]{stan-beta-correlates}
\caption{
  %Potential correlates of heavy-tailed dynamics.
  %(Top panels) Dots, thick lines, and thin lines represent fixed-effect estimates and their 50\% and 95\% confidence intervals for the covariates listed on the left: $\log \sigma$, $b$, $\lambda$, $\log$(lifespan), $\log$(data set length).
  Main effect posterior densitities for potential correlates of heavy-tailed dynamics.
  The response variable was Pr($\nu < 10$) (approximately the probability of heavy tails).
  The beta regression model was fit on a logit scale with random intercepts for taxonomic class, taxonomic order, and species.
  All covariates were standardized by subtracting their mean and dividing by twice their standard deviation. Therefore, the coefficient values represent the expected effect of a change in two standard deviations of the raw input variable on the logit-transformed response variable.
  %The beta distribution model (left panel) uses p$(\nu < 10)$ as the response (probability of heavy tails).
  %The binomial distribution model (right panel) uses p$(\nu < 10) > 0.5$ vs.\ p$(\nu < 10) \le 0.5$ as the response (heavy vs.\ not heavy tails).
  %The models were fitted as mixed effects models with beta or binomial distributions, logit links, and multilevel intercepts for taxonomic class and order nested within class using the \texttt{glmmADMB} \texttt{R} package \citep{fournier2012,glmmadmb}.
  %(Bottom panels) Model predictions for the strongest covariate --- time series length --- with all other predictors set to zero (their mean value when standardized). Dots represent observations for individual populations. Lines and shaded regions represent estimates and 50\% and 95\% confidence intervals.
}
    \label{fig:correlate-coefs}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.75\textwidth]{perc-prob-increase-with-n}
\caption{
Expected percent increase in the probability of detecting heavy tails, Pr($\nu < 10$), with an additional ten time steps of data.
These values are based on the curve shown in Fig.~\ref{fig:correlates}d, which is based on the coefficient in Fig.~\ref{fig:correlate-coefs}.
For example, we have about a 12\% increased probability of observing heavy tails in a time series that has 60 time steps vs.\ a time series with 50 time steps.
}
\label{fig:perc-inc-p}
\end{center}
\end{figure}

@@

\clearpage
\noindent
Stan code for the multilevel beta regression:
\begin{footnotesize}
\verbatiminput{betareg4.stan}
\end{footnotesize}
